---
title: "Assigment 2"
subtitle: ''
author: "Pablo E. Gutiérrez-Fonseca"
date: "`r format(Sys.time(), '%Y-%b-%d, %I:%M %p')`"
---
### Deadline / Fecha de entrega
**Deadline:** Monday, Sep 15, 2025; 6:00 PM.  
**Fecha de entrega:** Lunes, 15 de septiembre de 2025; 6:00 PM.

### Instructions / Instrucciones generales

- This assignment consists of 10 questions, each worth **10 points**.  
(Esta tarea consta de 10 preguntas, cada una vale **10 puntos**.)

- Write your answers in R code chunks whenever applicable.  
(Escribe tus respuestas en chunks de R cuando sea aplicable.)

- Make sure your code runs without errors. Include comments if necessary.  
(Asegúrate de que tu código se ejecute sin errores. Incluye comentarios si es necesario.)

- For questions about chunk options, experiment with the options and write down your observations.  
(Para preguntas sobre opciones de chunks, experimenta con ellas y escribe tus observaciones.)

- Submit the assignment on your personal webpages using the **HTML file** exported/knit from the `.Rmd`.  
Do not copy output screenshots unless requested.  
(Entrega la tarea en tus páginas personales usando el **archivo HTML** generado/kniteado desde el `.Rmd`.  
No copies capturas de pantalla del resultado a menos que se indique.)

<br><br><br>

#### **Using functions / Usando funciones**  
**Q1.** Convert the built-in **dataset BOD** to a list and print the **list**. **Q1.** Convierte el **conjunto de datos BOD** (predefinido en R) a una lista e imprime la **lista**.  

```{r}
BOD_df <- as.data.frame(BOD)# BOD (Biochemical Oxygen Demand) is pre-defined Vector in R.
BOD_df
```
<br><br>

#### **Converting matrix into DataFrame** 
**Q2.** Convert the following data frame into a data frame and print the matrix.
**Q2.** Convierte el siguiente data frame en un data frame e imprime la matriz.
```{r}
df1=data.frame(
  "Name"=c("a","b","c","d"),
  "Age"=c(12,34,23,67),
  "BMI"=c(22,12,10,11),
  "Country"=c("Malaysia","India","Spain","England"))
print(df1)
class(df1)
```
<br><br>

#### **Accessing 1st &2nd rows**
**Q3.** Create a subset of the `df` dataset that includes only the **first and second rows**. Print the subset..
**Q3.** Crea un subconjunto del `df` que incluya únicamente la **primera y la segunda fila**. Imprime el subconjunto.
```{r}
df=data.frame(
  A=c("Raju","Kumar","Chandru","Deepak","Raina", "Dhoni", "Hema", "Vandana", "Bela","Anand" ),
  B=c(12,34,23,67,23,43,23,38,34,12),
  C=c("M","M","M","M","M","M","F","F","F","M"),
  D=c("Malaysia","India","Spain","England", "Malaysia", "Malaysia", "India", "Egypt", "Italy", "Peru"))

names(df)<-c("Name", "Age", "Gender","Country" )
print(df)
class(df)
```
<br><br>

#### **Accessing 1st &2nd columns**
**Q4.** Create a subset of the df from Q3 that includes only the first and second columns. Print the subset.
**Q4.** Crea un subconjunto del df de la Q3 que incluya únicamente la primera y segunda columna. Imprime el subconjunto.
```{r}
# Your subset here
```
<br><br>

### **Sorting the dataframe**
**Q5.** Sort the df from Q3 by the `Age` column in ascending order. Print the sorted data frame.
**Q5.** Ordena el df de la Q3 por la columna `Age` en orden ascendente. Imprime el data frame ordenado.
```{r}
# Print the sorted subset
```
<br><br>

#### **Adding a new column**
**Q6.** Add a new column called **Score** to the df from Q3 with the following values: 85, 90, 78, 88, 90, 92, 80, 90, 87, 100. Print the updated data frame.
**Q6.** Agrega una nueva columna llamada **Score** al df de la Q3 con los siguientes valores: 85, 90, 78, 88, 90, 92, 80, 90, 87, 100. Imprime el data frame actualizado.
```{r}
# Your subset here
```
<br><br>

#### **Filtering rows with a condition**
**Q7.** From the `df` in Q3, create a subset where Age is greater than 30. Print the subset.
**Q7.** Del `df` de la Q3, crea un subconjunto donde Age sea mayor a 30. Imprime el subconjunto.
```{r}
# Your subset here
```
<br><br>

**Renaming columns**
**Q8.** Q8. Rename the columns of df from Q3 as **Person, Years, Gender, and Nation**. Print the updated data frame.
**Q8.** Renombra las columnas del df de la Q3 como Person, Years, Index y Nation. Imprime el data frame actualizado..
```{r}
# Your new column names here
```
<br><br>

#### **Selecting multiple conditions**
**Q9.** From the **df in Q2**, select rows where Age is greater than 20 and BMI is less than 15. Print the subset.
**Q9.** Del **df de la Q2**, selecciona las filas donde Age sea mayor a 20 y BMI menor a 15. Imprime el subconjunto.
```{r}
# Your subset here
```
<br><br>

#### **Summarizing a column**
**Q10.** Calculate the mean of the **Age** column from **df in Q2**. Print the result.
**Q10.** Calcula la media de la columna **Age** del **df de la Q2**. Imprime el resultado.
```{r}
# Your code here
```
<br><br>
