---
title: "Assignment 1"
subtitle: "BIOL6994/BIOL6997"
author: "Sebastian A. Cruz Romero"
date: "`r format(Sys.time(), '%Y-%b-%d, %I: %M %p')`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

## Q1. Assigning values

```{r q1}
height <- 175
height
```

## Q2. Simple math

```{r q2}
25 * 4 - 10
```

## Q3. Using functions

```{r q3}
sqrt(81)
```

## Q4. Character objects

```{r q4}
message <- "R is fun!"
message
```

## Q5. Object manipulation

```{r q5}
a <- 10
b <- 5
c <- a + b
c
```

## Q6. Logical comparison

```{r q6}
15 > 20
```

## Q7. Checking classes

```{r q7}
class("Hello World")
```

## Q8. Code chunk options

No `warning` argument
```{r q8step1}
log(-1)
```
`warning=FALSE`
```{r q8step2, warning=FALSE}
log(-1)
```

The warning appeared before adding `warning=FALSE`. After adding the option, the warning was hidden, but the result was still `NaN`.

## Q9. Code chunk options

No `message` argument
```{r q9step1}
# Load a package (may produce messages)
library(dplyr)
```
`message=FALSE`
```{r q9step2, message=FALSE}
# Load a package (may produce messages)
library(dplyr)
```

The package messages appeared before adding `message=FALSE`. After adding the option, the messages were hidden.

## Q10. Creating and analyzing a vector

```{r q10}
grades <- c(80, 90, 75, 95, 85)
average <- mean(grades)
average
```
