---
title: "Assigment 1"
subtitle: ''
author: "Pablo E. Gutiérrez-Fonseca"
date: "`r format(Sys.time(), '%Y-%b-%d, %I:%M %p')`"
---

### Instructions / Instrucciones generales

- This assignment consists of 10 questions, each worth **10 points**.  
(Esta tarea consta de 10 preguntas, cada una vale **10 puntos**.)

- Write your answers in R code chunks whenever applicable.  
(Escribe tus respuestas en chunks de R cuando sea aplicable.)

- Make sure your code runs without errors. Include comments if necessary.  
(Asegúrate de que tu código se ejecute sin errores. Incluye comentarios si es necesario.)

- For questions about chunk options, experiment with the options and write down your observations.  
(Para preguntas sobre opciones de chunks, experimenta con ellas y escribe tus observaciones.)

- Submit the assignment on your personal webpages using the **HTML file** exported/knit from the `.Rmd`.  
Do not copy output screenshots unless requested.  
(Entrega la tarea en tus páginas personales usando el **archivo HTML** generado/kniteado desde el `.Rmd`.  
No copies capturas de pantalla del resultado a menos que se indique.)

<br><br><br>

#### **Assigning values**
**Q1.** Create an object called height with the value 175. Print it. (Crea un objeto llamado height con el valor 175. Muéstralo en pantalla.)

<br><br>

#### **Simple math**
**Q2.** Calculate the result of 25 * 4 - 10. (Calcula el resultado de 25 * 4 - 10.)

<br><br>

####  **Using functions**
**Q3.** Use the function sqrt() to calculate the square root of 81. (Usa la función sqrt() para calcular la raíz cuadrada de 81.)

<br><br>

#### **Character objects**
**Q4.** Assign the phrase "R is fun!" to an object called message. Print it. (Asigna la frase "R is fun!" a un objeto llamado message. Muéstralo.)

<br><br>

#### **Object manipulation**
**Q5.** Create two objects a <- 10 and b <- 5. Calculate their sum and store it in c.
(Crea dos objetos a <- 10 y b <- 5. Calcula su suma y guárdala en c.)

<br><br>

#### **Logical comparison**
**Q6.** Test whether 15 is greater than 20.  (Verifica si 15 es mayor que 20.)  

<br><br>

#### **Checking classes**
**Q7.** Check the class of the object `"Hello World"`.  (Verifica la clase del objeto `"Hello World"`.)  

<br><br>

#### **Code chunk options**
**Q8.** Run the following code chunk:  
- Observe the warnings that appear. (Observa las advertencias que aparecen.)  
- Modify the chunk options **to suppress warnings** without changing the code itself. (Modifica las opciones del chunk para **apagar la advertencia** sin cambiar el código.)

```{r}
log(-1)
```

<br><br>

#### **Code chunk options**
**Q9.** Run the following code chunk:
- Observe the messages that appear. (Observa los mensajes que aparecen.)

- Modify the chunk options to suppress messages without changing the code itself.
(Modifica las opciones del chunk para apagar mensajes sin cambiar el código.)
```{r}
# Load a package (may produce messages)
library(dplyr)
```

<br><br>

#### **Help files**
**Q10.** Use R's help system to find out what the function `mean()` does. (Usa el sistema de ayuda de R para averiguar qué hace la función `mean()`.)

<br><br><br><br><br><br><br><br>


Q11
Suma 5+10
```{r}
5+10
```




