---
title: "Exam Practice"
subtitle: ''
author: "Pablo E. Gutiérrez-Fonseca"
date: "`r format(Sys.time(), '%Y-%b-%d, %I:%M %p')`"
---

### Load libraries

```{r message=F}
library(readxl)
library(dplyr)
```

### Practice data manupulation using Base R

The data in **cereal.csv** includes nutrition information collected from breakfast cereals. The data, which were provided by (Lock:2020?), were obtained from <http://nutritionresource.com/>.

-   The attributes included in the data are:
    -   name: Brand name of cereal
    -   company: Manufacturer (G=General Mills, K=Kellog’s or Q=Quaker)
    -   serving: Serving size (in cups)
    -   calories: Calories (per cup)
    -   fat: Fat (grams per cup)
    -   sodium: Sodium (mg per cup)
    -   carbs: Carbohydrates (grams per cup)
    -   fiber: Dietary Fiber (grams per cup)
    -   sugars: Sugars (grams per cup)
    -   protein: Protein (grams per cup)

```{r}
cereal <- read_excel("cereal.xlsx")
head(cereal)
```

### Five question (**2 pt each**):

1.  Extract rows 1 to 10 from the cereal data frame.
2.  Create a new data frame called Kelloggs that only contains rows where company == "K".
3.  From the cereal data, extract the names of all cereals that have more than 10 grams of sugar.
4.  Find the mean calories of all cereals in the dataset.
5.  Count how many cereals have fiber greater than or equal to 2 grams.

```{r, include=T}
# 1. Extract rows 1 to 10
cereal[1:10, ]

# 2. Create new data frame with only Kelloggs cereals
Kelloggs <- cereal[cereal$company == "K", ]
Kelloggs

# 3. Names of cereals with more than 10 grams of sugar
cereal$name[cereal$sugars > 10]

# 4. Mean calories of all cereals
mean(cereal$calories)

# 5. Count cereals with fiber >= 2 grams
length(cereal$fiber[cereal$fiber >= 2])
# (alternative: sum(cereal$fiber >= 2))
```

<br><br><br>

### Practice data manupulation using **dplyr** and **ktable**

The data in EquationCitations.csv includes information on the number of equations and citations for evolutionary biology publications. The data were obtained from Fawcett and Higginson (2012) and the online supplement (<https://www.pnas.org/doi/suppl/10.1073/pnas.1205259109/suppl_file/sd01.xlsx> ).

-   The attributes included in the data are:
-   journal: Journal in which the paper was published (The American Naturalist, Evolution, Proceedings of the Royal Society of London B: Biological Sciences)
-   authors: Names of authors
-   volume: Volume in which the paper was published
-   startpage: Starting page of publication
-   pages: Number of pages
-   equations: Number of equations in total
-   mainequations: Number of equations in main text
-   appequations: Number of equations in appendix
-   cites: Number of citations in total
-   selfcites: Number of citations by the authors themselves
-   othercites: Number of citations by other authors
-   theocites: Number of citations by theoretical papers
-   nontheocites: Number of citations by nontheoretical papers

**Details:** Fawcett and Higginson (2012) investigate how the number of equations in evolutionary biology papers affects citation counts. They found that papers with many mathematical equations receive fewer citations, especially from nontheoretical papers.

**Source:** Online supplements to Fawcett and Higginson (2012) <https://www.pnas.org/doi/suppl/10.1073/pnas.1205259109/suppl_file/sd01.xlsx>

**References:** Fawcett, T.W. and Higginson, A.D. (2012). Heavy Use of Equations Impedes Communication among Biologists. PNAS – Proceedings of the National Academy of Sciences of the United States of America, 109, 11735–11739. <doi:10.1073/pnas.1205259109>

```{r}
articles <- read.csv("EquationCitations.csv")
head(articles)
```

### Five question (**3 pts each**):

1.  Summarize by journal. Calculate the average number of equations per article in each journal. Display the results in a nicely formatted table. Hint: `group_by()` and `summarise()`

2.  **Find the top 5 articles** with the **highest number of total cites**. Show the authors, journal, and cites. Hint: `arrange(desc())`, `slice_head()`, and `select()`

3.  Create a new variable eq_ratio = mainequations / (mainequations + appequations). Then, compute the **mean** eq_ratio **per journal** and display in a table. Hint: `mutate()`, `group_by()`, and `summarise()`

4.  Theoretical vs. non-theoretical citations. For each **journal**, calculate the **average proportion of theoretical citations**. Hint: `mutate()`, `group_by()`, `summarise()`

5.  **Filter and summarize short articles.** Filter articles with **fewer than 10 pages**. For these, calculate the **average number of equations** and the **average** number of citations. Hint: `filter()` and `summarise()`

```{r, include= FALSE}
library(dplyr)
library(knitr)

# 1. Summarize by journal: average number of equations
articles %>%
  group_by(journal) %>%
  summarise(avg_equations = mean(equations, na.rm = TRUE)) 

# 2. Top 5 articles with highest total cites
articles %>%
  arrange(desc(cites)) %>%
  slice_head(n = 5) %>%
  select(authors, journal, cites)

# 3. Create eq_ratio and compute mean per journal
articles %>%
  mutate(eq_ratio = mainequations / (mainequations + appequations)) %>%
  group_by(journal) %>%
  summarise(mean_eq_ratio = mean(eq_ratio, na.rm = TRUE))

# 4. Proportion of theoretical citations per journal (simpler version)
articles %>%
  mutate(theo_ratio = theocites / cites) %>%
  group_by(journal) %>%
  summarise(avg_theo = mean(theo_ratio, na.rm = TRUE))

# 5. Filter short articles (<10 pages) and summarize
articles %>%
  filter(pages < 10) %>%
  summarise(
    avg_equations = mean(equations, na.rm = TRUE), 
    avg_cites = mean(cites, na.rm = TRUE)
  ) 

```
<br><br><br>

### Practice data manupulation using **dplyr** and **ktable**

The data in EquationCitations.csv includes information on the number of equations and citations for evolutionary biology publications. The data were obtained from Fawcett and Higginson (2012) and the online supplement (<https://www.pnas.org/doi/suppl/10.1073/pnas.1205259109/suppl_file/sd01.xlsx> ).

-   The attributes included in the data are:
-   journal: Journal in which the paper was published (The American Naturalist, Evolution, Proceedings of the Royal Society of London B: Biological Sciences)
-   authors: Names of authors
-   volume: Volume in which the paper was published
-   startpage: Starting page of publication
-   pages: Number of pages
-   equations: Number of equations in total
-   mainequations: Number of equations in main text
-   appequations: Number of equations in appendix
-   cites: Number of citations in total
-   selfcites: Number of citations by the authors themselves
-   othercites: Number of citations by other authors
-   theocites: Number of citations by theoretical papers
-   nontheocites: Number of citations by nontheoretical papers

**Details:** Fawcett and Higginson (2012) investigate how the number of equations in evolutionary biology papers affects citation counts. They found that papers with many mathematical equations receive fewer citations, especially from nontheoretical papers.

**Source:** Online supplements to Fawcett and Higginson (2012) <https://www.pnas.org/doi/suppl/10.1073/pnas.1205259109/suppl_file/sd01.xlsx>

**References:** Fawcett, T.W. and Higginson, A.D. (2012). Heavy Use of Equations Impedes Communication among Biologists. PNAS – Proceedings of the National Academy of Sciences of the United States of America, 109, 11735–11739. <doi:10.1073/pnas.1205259109>

```{r}
articles <- read.csv("EquationCitations.csv")
head(articles)
```

### Question (**10 pts**):

The journal Proceedings of the National Academy of Sciences has asked you to revisit the data from Fawcett and Higginson (2012) on the use of equations in evolutionary biology papers. They are particularly interested in two things:

1. On average, how many equations per article are published in each journal? (Hint: use group_by() and summarise())

2. For each journal, calculate the average proportion of theoretical citations relative to total citations. (Hint: create a new column with mutate(theo_ratio = theocites / cites), then group_by() and summarise())

```{r}
library(dplyr)

# 1. Average number of equations per article in each journal
articles %>%
  group_by(journal) %>%
  summarise(avg_equations = 
              mean(equations, na.rm = TRUE))

# 2. Average proportion of theoretical citations relative to total citations, per journal
articles %>%
  mutate(theo_ratio = theocites / cites) %>%
  group_by(journal) %>%
  summarise(avg_theo_ratio = mean(theo_ratio, na.rm = TRUE))
 
```
<br><br>

#### Potential findings:<br> 
My findings showed that, on average, the number of equations per article varied among journals. In particular, The American Naturalist, Evolution, and Proceedings of the Royal Society B showed differences, with the average ranging from ______ (Am. Nat.) to ______ (Evolution) to ______ (Proc. B) equations per article. In terms of citations, the proportion of theoretical citations relative to total citations also varied by journal. On average, articles in The American Naturalist had a mean theoretical citation ratio of _____, those in Evolution had _____, and those in Proceedings B had _____.
<br><br><br><br>