---
title: "Study Guide: Base R "
subtitle: ''
author: "Pablo E. Gutiérrez-Fonseca"
date: "`r format(Sys.time(), '%Y-%b-%d, %I:%M %p')`"
---

<br><br>

#### 1. `help()` Get documentation for a specific function.
```{r eval= F}
  
help(mean)       # Open help page for mean()
?mean            # Equivalent shortcut
```
<br><br>

#### 2. `c()` Combine values into a vector.
```{r warning=FALSE, message=FALSE}
x = c(8, 6, 7, 5, 3, 0, 9)
x
```
<br><br>

#### 3. `data()` Load built-in datasets.
```{r}
data()           # List available datasets
data(iris)       # Load the iris dataset
head(iris)       # Show first rows
```
<br><br>

#### 4. `View()` Open a dataset in spreadsheet format.
```{r eval=FALSE}
View() # View dataset in a spreadsheet-type format
```
<br><br>

#### 5. `str()` Display the internal structure of an object.
```{r}
str(iris)
```
<br><br>

#### 6. `read.csv()`, `read.table()` Load external data files into R.
```{r eval=FALSE}
mydata <- read.csv("file.csv")
mydata <- read.table("file.txt", header=TRUE)
```
<br><br>

#### 7. `library()` Load add-on packages.
```{r eval=FALSE}
library(ggplot2)   # Load the ggplot2 package
```
<br><br>

#### 8. `dim()` Check the dimensions (rows × columns) of a dataset.
```{r}
dim(iris)
```
<br><br>

#### 9. `length()` Check the length of a vector.
```{r }
length(x)
```
<br><br>

#### 10. `mean()` Calculate the arithmetic mean.
```{r}
y <- iris$Petal.Length
mean(y)
```
<br><br>

#### 11. `summary()` Get descriptive statistics.
```{r }
summary(y)
```
<br><br>

#### 12. Subsetting with logical conditions
```{r}
# Extract rows 1 to 10 from a dataset
iris[1:10, ]

# Subset rows where Sepal.Length is greater than 7
subset(iris, Sepal.Length > 7)

# Extract only the Species names where Petal.Width is less than 0.2
iris$Species[iris$Petal.Width < 0.2]
```


## Creating Objects in R

- R works with different types of objects:
```{r}
# Vector
v <- c(1, 2, 3, 4)

# Matrix
m <- matrix(data=5, nrow=2, ncol=2)

# Data frame
df <- data.frame(Name=c("Ana","Luis"), Age=c(23,30))

# List
lst <- list(numbers=v, matrix=m, people=df)

v
m
df
lst
```

