---
title: "Study Guide: dplyr for Data Manipulation"
subtitle: ''
author: "Pablo E. Gutiérrez-Fonseca"
date: "`r format(Sys.time(), '%Y-%b-%d, %I:%M %p')`"
---
<br><br>

#### 1. library() Load dplyr
Required for all dplyr functions.
```{r, message=FALSE}
library(dplyr)
```
<br><br>

#### 2. `%>%` Pipe operator
Chains commands.
Sends output of one function to the next. Essential for readable workflows.
```{r eval=FALSE}
data %>%
  filter(condition) %>%
  group_by(group_var) %>%
  summarise(mean_value = mean(var))
```
<br><br>

#### 3. `filter()` Subset rows
Basic filtering

```{r eval=FALSE}
data %>% filter(variable > 5)
```
<br><br>


#### 4. Multiple conditions
Use &, |, and ! for logical combinations.

```{r eval=FALSE}
data %>% filter(variable > 5 & other_var == "A")
```
<br><br>

#### 4. `select()` Pick columns
```{r, eval=FALSE}
data %>% select(col1, col2, col3)

# Helpers
data %>% select(starts_with("eq"))   # all columns starting with 'eq'
data %>% select(col1, col2, everything()) # keeps all columns, but moves col1 and col2 first
```
<br><br>

#### 5. mutate() Create or transform columns
Create new column
```{r, eval=FALSE}
data %>% mutate(new_col = var1 / var2)
```

#### Multiple columns
Create multiple columns at once.  Very useful to compute ratios or normalized values.
```{r, eval=FALSE}
data %>% mutate(
  ratio = a / b,
  log_c = log(c)
)
```
<br><br>

#### 6. `group_by()` Group data
Must use before summarise() to calculate per-group statistics.
```{r, eval=FALSE}
data %>% group_by(group_var)
```
<br><br>

#### 7. `summarise()` Summarize data
Commonly used with group_by() for per-group summaries.

```{r, eval=FALSE}
data %>%
  group_by(group_var) %>% # looks at the values in the column group_var and splits the rows into groups based on those values 
  summarise(mean_var = mean(var, na.rm=TRUE), # average of 'var' within each group
            sum_var = sum(var, na.rm=TRUE), # total of 'var' within each group
            count = n()) # number of rows in the group
```
<br><br>

#### 8. arrange() Sort rows
```{r eval=F}
# Ascending
data %>% arrange(var)

# Descending
data %>% arrange(desc(var))

```
<br><br>

#### 9. slice_head() and slice_tail() Select top/bottom rows
Works well after `arrange()` to get top or bottom items.
```{r eval=FALSE}
data %>% slice_head(n=5)   # first 5 rows
data %>% slice_tail(n=5)   # last 5 rows
```
<br><br>

#### 10. Combining tools: Typical workflow
```{r, eval=FALSE}
data %>%
  filter(condition) %>%           # Step 1: filter rows
  mutate(new_col = var1/var2) %>%# Step 2: create new variable
  group_by(group_var) %>%         # Step 3: group by a factor
  summarise(avg_val = mean(new_col, na.rm=TRUE)) %>%  # Step 4: summarise
  arrange(desc(avg_val))          # Step 5: sort descending
```
<br><br><br><br>
